Towards Verifying the Bitcoin-S Library

### 2020-07-24 paper-short/main.tex
work on final version

We are happy to see some friendly green verifier output. However,
apart from the bugs we found, the main conclusion from this work is
that we had to non-trivially transform even a very small portion of
the code in order to verify it. At the moment, it is unrealistic to
routinely formally verify properties as part of the Bitcoin-S
development process. However, Stainless development has already
progressed (e.g.\ type members are supported in recent versions) and
continues to do so (e.g. self-referential type bounds are on the
roadmap). Some missing features that we identified are presumably very
easy to support, like the message parameter in the require
function. Some other features presumably require more substantial
work, like bitwise operations on integer types.

On the other hand, Bitcoin-S uses features that might not be supported
even by future Stainless versions, such as calls to Java code. Here
the Bitcoin-S code can hopefully be adapted to accommodate formal
verification.

From our results we conclude that formal verification of bitcoin
libraries in general and Bitcoin-S in particular is a worthwhile
endeavour. We are looking forward to verifying more substantial parts
of the code in future work.

>>>>

We are happy to see some friendly green verifier output. However,
apart from the bugs we found, the main conclusion of this work is that
we had to non-trivially transform even a very small portion of the
code (70 lines) in order to verify it. And that was true even though
the code was purely functional to begin with. At the moment, it is
probably unrealistic to routinely formally verify properties as part
of the Bitcoin-S development process. However, Stainless development
has already progressed (e.g.\ type members are supported in recent
versions) and continues to do so (e.g. self-referential type bounds
are on the roadmap). Some missing features that we identified are
presumably very easy to support, like the message parameter in the
require function. Some other features presumably require more
substantial work, like bitwise operations on integer types.

On the other hand, Bitcoin-S uses features that might not be supported
even by future Stainless versions, such as calls to Java code.

Given our experience, the best route towards integrating verification
into the Bitcoin-S development process would be to re-implement parts
of the library in Pure Scala. We would split the library into a
\emph{verified} and \emph{non-verified} part, and use Stainless only
on the \emph{verified} part. It is then both a technical but also a
political question how much code, if any, can be moved to the
\emph{verified} part. That is an interesting direction for future
work.
